NumPy---Array Programming in Python

### 2020-01-28 body.tex
Minor update to Scientific Python ecosystem.
Most important change is to talk about numpy.distutils and f2py
rather than minimum versions; the latter really is a small detail.

SciPy and Matplotlib are tightly coupled with NumPy---in terms of
history, development, and use.
% to expand the suite of standard algorithms and
% provide basic plotting capabilities.
SciPy provides fundamental algorithms for scientific computing
including mathematical, scientific, and engineering routines.
Matplotlib generates publication-ready figures and visualizations.
The combination of NumPy, SciPy, and Matplotlib together with
an advanced interactive environment like IPython \cite{perez2007ipython}
or Jupyter \cite{Kluyver:2016aa}
provides a complete foundation for array-based scientific computing.
The scientific Python ecosystem (Fig.~\ref{fig:ecosystem}) builds on top of
this foundation to provide several, widely used \emph{technique-specific}
libraries\cite{pedregosa2011scikit,vanderwalt2014scikit,SciPyProceedings_11},
that in turn underlay numerous \emph{domain-specific} projects
\cite{astropy:2013,astropy:2018,cock2009biopython,millman2007analysis,sunpy2015,2018EGUGA..2012146H}.
NumPy, as the focal point for the ecosystem of array-aware libraries,
sets documentation standards, provides array testing infrastructure,
and defines minimum version dependencies.

@@

For example, the \code{eht-imaging} library developed by the Event Horizon Telescope 
collaboration relies on many components of the scientific Python ecosystem.
NumPy arrays are used to store and manipulate numerical data at every step
in the processing chain: from raw data through calibration and image 
reconstruction.
SciPy is used to provide tools used in general image processing such as 
filtering and image alignment, while scikit-image---a popular image processing
library that extends SciPy---provides higher-level functionality such as
edge filters and Hough transforms.
Other components of SciPy are used throughout the library, such as the
\code{scipy.optimize} subpackage for handling general optimization tasks.
NetworkX \cite{SciPyProceedings_11}---a popular package for complex
network analysis---is used for analysis of consistency among image
comparisons.
The community developed astronomy library, Astropy \cite{astropy:2013, astropy:2018},
is used throughout \code{eht-imaging},
primarily to handle standard astronomical file formats and time/coordinate
transformations.
Matplotlib is
used for visualizing data throughout the analysis pipeline,
including the generation of the final image of the black hole.
% N.B. - most of this comes from introspecting the Image class defined in
% ehtim/image.py and grepping the dependencies in requirements.txt

@@

Similar examples can be found in physics, earth science, biology,
and more.
Researchers in these fields depend on NumPy being correct, reliable,
and repeatable.
The rich ecosystem of tools building on NumPy and the larger community
of scientists and programmers using those tools mean that NumPy has a
responsibility to maintain considerable stability.
At times, this conflicts with the need to innovate;
resolving it requires balancing trade-offs.
% our focus is ...
% NumPy purpose engages with this problem by ...
% [this also defines the scope of NumPy]
This need creates interesting challenges for NumPy development as the
computational environment scientists are analyzing their data on continues
to evolve and change.
% this leads to a proliferation of arrays

>>>>

SciPy and Matplotlib are tightly coupled with NumPy---in terms of
history, development, and use.
% to expand the suite of standard algorithms and
% provide basic plotting capabilities.
SciPy provides fundamental algorithms for scientific computing
including mathematical, scientific, and engineering routines.
Matplotlib generates publication-ready figures and visualizations.
The combination of NumPy, SciPy, and Matplotlib together with
an advanced interactive environment like IPython \cite{perez2007ipython}
or Jupyter \cite{Kluyver:2016aa}
provides a complete foundation for array-based scientific computing.
The scientific Python ecosystem (Fig.~\ref{fig:ecosystem}) builds on top of
this foundation to provide several, widely used \emph{technique-specific}
libraries\cite{pedregosa2011scikit,vanderwalt2014scikit,SciPyProceedings_11},
that in turn underlay numerous \emph{domain-specific} projects
\cite{astropy:2013,astropy:2018,cock2009biopython,millman2007analysis,sunpy2015,2018EGUGA..2012146H}.
NumPy, as the focal point for the ecosystem of array-aware libraries,
sets documentation standards, and provides array testing infrastructure
and build support for integrating with Fortran code and using specialized
compilers and computing platforms.

@@

For example, the \code{eht-imaging} library developed by the Event Horizon Telescope 
collaboration relies on many components of the scientific Python ecosystem.
NumPy arrays are used to store and manipulate numerical data at every step
in the processing chain: from raw data through calibration and image 
reconstruction.
SciPy is used to provide tools used in general image processing such as 
filtering and image alignment, while scikit-image---a popular image processing
library that extends SciPy---provides higher-level functionality such as
edge filters and Hough transforms.
Other components of SciPy are used throughout the library, such as the
\code{scipy.optimize} module for handling general optimization tasks.
NetworkX \cite{SciPyProceedings_11}---a popular package for complex
network analysis---is used for analysis of consistency among image
comparisons.
The community-developed astronomy library, Astropy \cite{astropy:2013, astropy:2018},
is used throughout \code{eht-imaging}, primarily to handle standard
astronomical file formats and time/coordinate transformations.
Matplotlib is used for visualizing data throughout the analysis pipeline,
including the generation of the final image of the black hole.
% N.B. - most of this comes from introspecting the Image class defined in
% ehtim/image.py and grepping the dependencies in requirements.txt

@@

Similar examples can be found in physics, earth science, biology,
and more.
Researchers in these fields depend on NumPy being correct, reliable,
and repeatable.
The rich ecosystem of tools building on NumPy and the larger community
of scientists and programmers using those tools mean that NumPy has a
responsibility to maintain considerable stability.
At times, this conflicts with the need to innovate;
resolving it requires balancing trade-offs.
% our focus is ...
% NumPy purpose engages with this problem by ...
% [this also defines the scope of NumPy]
This need creates interesting challenges for NumPy development as the
computational environment scientists are analyzing their data in continues
to evolve and change.
% this leads to a proliferation of arrays
